```{r Setup, include=FALSE}
library(MotifBinner)
```

# Investigation of various approach to constructing a consensus string

## Overview

Three important steps needs to be optimized:
* mislabel classification
* alignment
* consensus string construction

This document investigates all three steps simultaneously by a simulation
process:
* Given an input string and a read error profile
* Potentatially given a contamination string
* Simulate reads from both strings and form a bin
* Use mislabel classification to throw out outliers
* Align the remaining sequences
* Construct a final sequence
* Compare the reconstructed sequence to the input sequence with edit distance

We need very clear terminology to keep track of the different aspects of the
simulations:
* All the parameters that goes into the three steps are called a 'setup'
* A input sequence, number of reads, contamination sequence, number of
  contamination reads and the error profile is called a 'scenario' or 'test
  scenario'.
* A test scenario, setup and a seed is called a 'case' or 'test case'.

## The setups

```{r}
setups <- list()
setups[['base']] <- list(classification_technique = 'infovar_balance', 
                         classification_params = list(threshold = 1), 
                         alignment_technique = 'muscle', 
                         alignment_params = list(), 
                         consensus_technique = 'Biostrings::consensusString', 
                         consensus_params = list())
```

## The error profiles

```{r}
err_profiles <- list()
x <- list(read_len = 1000,
          A_err_rates = list('A' = 0.997, 'C' = 0.001, 'G' = 0.001, 'T' = 0.001),
          C_err_rates = list('C' = 0.997, 'A' = 0.001, 'G' = 0.001, 'T' = 0.001),
          G_err_rates = list('G' = 0.997, 'C' = 0.001, 'A' = 0.001, 'T' = 0.001),
          T_err_rates = list('T' = 0.997, 'C' = 0.001, 'G' = 0.001, 'A' = 0.001))
params <- list()
params[['params']] <- x
params[['technique']] <- 'uniform'
err_profiles[['unif_1_1000']] <- do.call(gen_error_profile, params)
```

## The test scenarios

```{r}
scenarios <- list()
scenarios[['stupid_easy']] <- list(ref_seq = paste(rep('A', 50), collapse = ""), 
                                   n_reads = 10, 
                                   error_rates = err_profiles[['unif_1_1000']], 
                                   contam_seq = NULL, 
                                   n_contam = 0)
```

## Results

```{r}
params <- scenarios[['stupid_easy']]
params[['seed']] <- 1

test_bin <- do.call( gen_and_contaminate_reads, params)

params <- setups[['base']]
params$test_bin <- test_bin
result <- do.call(score_consensus, params)
print(result$edit_dist)
```
